---
title: "Convert (Reclassify) SpeciesNet Taxonomic Category"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Convert (Reclassify) SpeciesNet Taxonomic Category}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette illustrates how a user can update the categories provided in a SpeciesNet JSON file to better suit their workflow/project. This example will show reclassifying names into fewer categories with the intent of making it easier to go by species. For this, we'll use a subset of 500 images from the [ENA24-detection](https://lila.science/datasets/ena24detection). This dataset comes from eastern North America, so we'll focus on cutting out any species that are not present in this area and combining classifications that can only really be one thing into a single category.

For example, converting all detections tagged as the cervidae family and *Odocoileus* genus to white-tailed deer (*Odocoileus virginianus*). If you run the general SpeciesNet model and not a country/state specific one (like I have), you will probably want to reassign things that are taxonomically different to a species found in your area, like muntjac deer and other Eurasian cervids to white-tailed dear for North American images.The inverse could also be done if SpeciesNet is classifying things higher taxonomically than you want, such as tagging detections as *peromyscus* when you can't tell anything more than the animal is some kind of rodent.

```{r setup}
library(jsonlite, camRa)

#Turn off validation
options("camRa.validate_json" = FALSE)
```

First, we'll load in the data we're using. The subset file names are provided as vector, while SpeciesNet data is stored as a JSON file. The subset can be called at anytime after the package is loaded, but the JSON file needs to be called with `system.file()`.

```{r}
#Copy the path to the JSON file
json_file <- system.file(
  "extdata", 
  "ena24subset_SpecNet_recognition.json", 
  package = "camRa"
)
```

JSON files for both SpeciesNet and MegaDetector map categories using numbers instead of rewriting the longer species name in each detection. This map is stored as one of the top-level or root keys, "classification_categories". We want to pull out a list of the species names to build our table defining what names get changed to what (our "map"), along with the descriptions used for these names incase you also want to change those.

These values can be obtained using two different methods. You can load the JSON file and index with `$` or use `megadet_get_info()` if you want to avoid reading in the json file to an object or avoid importing `jsonlite`, since it'll do it for you. Regardless of method, since species names are the values in a named list, you must use `unname()` and `unlist()` to extract them.

Note: `megadet_get_info()` will also accept `json_data` instead of the file path. Use this if you prefer having the function to index but will use the JSON data multiple times in your script to avoid reloading it everytime you use a `camRa` function.

```{r}
#Read in JSON data
json_data <- jsonlite::read_json(json_file)

#Get values directly (method 1)
species_names <- json_data$classification_categories |> 
  unname() |> unlist()

#Get values using index function (method 2)
species_names <- camRa::megadet_get_info(
  json_file, 
  key = "classification_categories"
) |> unname() |> unlist()
```

This process is repeated for the descriptions attached to each species, which are stored under a different key but use the same format.

```{r}
#Get values directly (method 1)
species_desc <- json_data$classification_category_descriptions |> 
  unname() |> unlist()

#Get values using index function (method 2)
species_desc <- camRa::megadet_get_info(
  json_file, 
  key = "classification_category_descriptions"
) |> unname() |> unlist()
```

Next, we want to build a table from this information and save it somewhere so it can be more easily edited. We'll just use a CSV since that's easy and works fine for small files. After this file is written, you can fill in the new_species and (optionally) the new_description columns with the values you want. Only the first description value associated with each species will be used, so ensure these are consistent. I.e., if you change both "cervidae" and white-tailed deer" to "Deer", both of these rows should have the same description..

Values mapped from the old to new species can be matched in a many-to-one relationship, so the multiple species can be remapped to a single one. The new values also do not have to exist in the old species column. Feel free to abbreviate names (like just "raccoon" instead of the default "northern raccoon") or change them entirely ("canid family" to "probably just a coyote").

Be careful re-running the cell below, as it can overwrite your hard work filling out the CSV. I suggest renaming the file to prevent this before you start editing.

```{r}
#Create table as list of vectors
##No need to convert to a data frame here
species_map <- list(
  "old_species" = species_names,
  "old_description" = species_desc,
  "new_species" = rep("", length(species_names)),
  "new_description" = rep("", length(species_desc))
)

#Write to CSV
write.csv(species_map, "species_map.csv", row.names = FALSE)
```

After the CSV is filled in with new species values, we'll read it back in and give the columns to the reclassification function. We only care about the new and old species columns, and optionally the new description column. If you'll be repeating this with many JSON files and reusing this file, it may be good to add extra information in an additional column, such as why you reclassified things the way you did.

Below, we're reading in a stored CSV file from he package instead of reading in the same file we output (or, a renamed copy). Generally, you'll want to change this to be a typical file path. You'll see in this file, a new column called "reclass_rational" was added, which contains info on why certain things were changed.

```{r}
#Copy the path to the CSV file
csv_file <- system.file(
  "extdata", 
  "species_map_ena24subset.csv",
  package = "camRa"
)

#Read in CSV
species_map <- read.csv(csv_file)

#Show table
##Remove old description for view since its long text
knitr::kable(head(species_map[, !names(species_map) %in% "old_description"]))

```

With our table, we just need to give the columns to `camRa::specnet_reclassify()` and tell it where to save the new file.

The CSV used here has information in the new_description column, but if you do not want to update these, usually you should use `NULL` for this parameter or remove it from the function call, as you may get strange behavior when multiple `NA`s are passed. Alternatively, convert all `NA`s to empty character strings.

```{r}
#Reclassify JSON
new_json_data <- camRa::specnet_reclassify(
  json = json_data, 
  values_from = species_map$old_species,
  values_to = species_map$new_species,
  values_description = species_map$new_description,
  file = "reclass_ena24subset_MegaDet_recognition.json"
)
```

Now you have an updated file, but how can you tell if it's correct? With the way SpeciesNet data is written, the JSON file just has numbers in the detections and you have to go check the classification_categories key for updated values then go searching through the image data to see what the numbers match up as... Instead, let's flatten the JSON files and compare the two columns using `megadet_flatten()`. By default, this function will map species names onto the image data to avoid just looking at numbers, but I've included `map_names = TRUE` anyways.

We don't care about the other fields this gives us, so we'll just grab the "classification_category" field from the tables the function produces instead of saving the entire table to an object and print them out together.

After running this, you can see that the two json files now have different values that match what was defined in our map file.

```{r}
#Get JSON info as tables, then grab column
species_old <- camRa::megadet_flatten(
  json = json_data,
  map_names = TRUE
)$classification_category

species_new <- camRa::megadet_flatten(
  json = new_json_data,
  map_names = TRUE
)$classification_category

#Print random 20 value pairs
##Use random because sequential are probably mostly the same sp.
set.seed(98632)
rand_index <- sample(1:500, size = 20)

for (i in rand_index) {
  cat(
    "Old Taxon: ", species_old[[i]], 
    ", New Taxon: ", species_new[[i]], 
    "\n",  sep = ""
  )
}
```
