---
title: "Style Guidelines/Conventions for this Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Style Guidelines/Conventions for this Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(camRa)
```

This page is largely for my own reference, but has been added as a vignette as it has some information that may be useful to users or contributers. When writing this package, I have some conventions that make sense to me, but may sometimes conflict with what others do; sometimes, these are just general good practices as well. Some things I've adapted since learning why I shouldn't be doing them, like using `NA` as a default argument (bad!, should use `NULL`) and added them here.

# Limiting Package Imports

When writing code, as a general rule I want to avoid adding in any extra imports that I don't absolutely need to reduce the possibility of incompatibility in the future and reduce what users must install to use this package. For that reason, some utility packages such as `dplyr` haven't been used even though they very well could be. Unless something cannot be done within reason without a package (like reading in JSON data with `jsonlite`), it will not be considered.

# Handling Files/Imports for the User

When using this package for basic data exploration or just testing things out, users should not have to import any other packages. For this reason, when you use JSON or image data, either paths or the read file as an R object will be accepted. Along with this, every function that returns data formatted as a file type that can only reasonably be saved using another package includes a `file` argument to specify where it should be saved. To prevent users from accidently overwriting data, these functions also include an `overwrite` parameter.

Some functions do not have a `file` argument, while similar ones do. Most MegaDetector functions return JSON formatted data and use `file`, but `megadet_get_counts()` returns a table that can be saved with base `write.csv()` or similar functins, so it doesn't.

# Function Naming

Functions all use snake case ("first_second_third") and are limited to three words (i.e., up to two underscores. With snake case, all functions use undercase with the exception of when acronyms (like "LILA" for the Living Library of Alexandria) are in the function. Limiting the number of words in a function name will make it less of a pain to use and if possible names should be made shorter like not using `megadet_filter_json()` and instead using `megadet_filter()`.

Functions typically use prefixes that align with their general use case, which is also generally what .R file they are defined in. For MegaDetector and SpeciesNet functions, the approach has been to use "megadet\_" for functions that can apply to files with information with both and "specnet\_" for any that only apply to files that have SpeciesNet data. Most image functions will be prefaced with "img\_" which was done to prevent conflicts with `magick` functions (they use "image\_").

# Globals/Settings

This package has some global settings to prevent needing to rewrite the same argument over and over. As of writing, the only global setting is `camRa.validate_json` which determines if the `json` argument in functions should be validated as a properly formatted object (only applies if the input is not a path to the file). `jsonlite` does not give its outputs a unique class, so this is done by trying to convert it to real JSON. For large files covering hundreds of thousands of images, validation will slow things down immensely but it is important to handle strange errors that could happen if you give `json` something it doesn't expect. Users will likely want to turn this off if they know what they're doing, so the setting is editable using `options()`. This can also be done by function using the `validate_json` argument (default = `TRUE`).

# On the Use of Quotes

When writing in R, both single and double quotes (`'text'` and `"text"`) are acceptable. As a personal preference, I use single quotes for everything. The reasoning for this is just practicality, my keyboard has both on the same key and double quotes is the one that needs you to use the shift key (annoying!). I've found it also makes it easier to write error strings, since I often want to say X argument must be "x", "y", or "z", and using double quotes as text is only possible usually with single quotes wrapping the text. To keep consistent, this is a rule for the entire package, including potential contributions from others. The only exception is in error strings as mentioned before and vignettes. Most people from my experience use double quotes, so vignettes will use them. It matters more than the code that most users will never actually neded to look at (or at least, I hope).

# Default Function Values

`NULL` should be the default when there isn't a value that belongs there. In the past, I've used mainly `NA`, but for consistency am enforcing `NULL` as the proper value.
